### 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/coverletter_v3.tex
Added the panlex paper to the repo

Again we thank the reviewers and editors for their constructive feedback on our submission.

Overall, we performed the following changes:
- Section 1 (Introduction): The description of PanLex was shortened. We emphasized the role of PanLex, as we see it, in the Web of Data.
- Section 2 (Triplification): We added a section on vocabulary reuse and added according depicitions. Added a table with approvers per license type.
- Section 4 (Publication): Improved descriptions of the published resources, especially that downloads are tagged with a timestamp of their creation. 
- Section 5 (Benefits and Usage Scenarios): Pointed to the web site for SPARQL query examples due to space reasons. Added cross language information retrieval and DBpedia spotlight (entity disambiguation) as potential applications.
- Section 6 (Related Work): Minor explanation of the relationship between wordnets and PanLex.
- Section 7 (Conclusions and Future Work): Improved descriptions of directions for future work, notably that descriptions of datasets within PanLex could be modeled with VoID.